---
title: "Tidy Tuesday API"
subtitle: "Build and Deploy a Plumber API"  
author: "Andrew Bates"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo(primary_color = "#1F4257", secondary_color = "#F97B64")
```

## Tidy Tuesday Project

> A weekly data project aimed at the R ecosystem. 

<br/>
<br/>

Run by **Tom Mock** and the **R4DS Online Learning Community**

<br/>

A new data set is published every week. Post content on Twitter with
 **#TidyTuesday**.

---
## Plumber

> Turn your R code into a web API

<br/>

```{r plumber-add, eval = FALSE}
#* Return the sum of two numbers
#* @param a The first number to add
#* @param b The second number to add
#* @post /sum
function(a, b) {
  as.numeric(a) + as.numeric(b)
}
```

---
class: inverse, center, middle

## Tidy Tuesday + Plumber = Tidy Tuesday API

---
class: center, middle

## What is an API?


???

machine to machine communication over the internet

show how to access TT API in browser


---
## Why?

contrib to community
data sets for non-R users
resume bullet point
fun!

---
## How

- Data sets stored in [mongo DB](https://www.mongodb.com/) database
- API fetches/sends results
- API hosted on [Digital Ocean](https://www.digitalocean.com/)

???

shout out to plumberDeploy!

---
## Mongo DB

- Document database (NoSQL)
- More flexible structure vs. tables
- Connect from R via [`{mongolite}`](https://jeroen.github.io/mongolite/)

--

.pull-left[
**SQL**

- database
- table
- record/row
]

.pull-right[
**mongo DB**

- database
- collection
- document
]


---
## API

show code
how does it work
what are the endpoints
etc.

---
## Digital Ocean + plumberDeploy

what is digital ocean
what is a droplet
what is plumberDeploy
how does it help

---
class: inverse, center, middle

## Problems

problems i ran into
maybe think of a better word (issues?)

---
## Mongo DB

oh the troubles i've seen...

data too big
spent **a lot** of time searching for a solution (Azure cosmos db, Amazon
 dynamo? db, etc., etc.)
eventually used mongo large file system (what's it called?, how is it different?)

---
## API

cors
how to handle db connections?
mongolite

---
## Digital Ocean

plumberDeploy
what are some of the issues I came across?
how did i overcome them?

---
class: inverse, center, middle
## title to break things up

---
## What I learned

stress: i didn't know much going into this. only that plumber looked simple
 enough and that it was probably possible

mongodb, document databases
plumber
deploying
digital ocean
domain name

---
## Why I'm sharing

maybe you want to make a plumber api
show that it's not all rainbows and unicorns
hopefully help motivate to do side projects, especially for newer folks
yes, it may be hard, but that doesn't mean you shouldn't try

---
## Resources

links to tidy tuesday
links to tt api
links to some docs (i.e. plumber, plumberDeploy)
link to this talk (have at beginning too. title slide probably)

---
class: center, middle

# Thanks!

Slides created via the R packages:

[**xaringan**](https://github.com/yihui/xaringan)<br>
[gadenbuie/xaringanthemer](https://github.com/gadenbuie/xaringanthemer)

The chakra comes from [remark.js](https://remarkjs.com), [**knitr**](http://yihui.name/knitr), and [R Markdown](https://rmarkdown.rstudio.com).
