---
title: "Tidy Tuesday API"
subtitle: "Build and Deploy a Plumber API"  
author: "Andrew Bates"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo(primary_color = "#1F4257", secondary_color = "#F97B64")
```

## Tidy Tuesday Project

> A weekly data project aimed at the R ecosystem. 

<br/>
<br/>

Run by **Tom Mock** and the **R4DS Online Learning Community**

<br/>

A new data set is published every week. Post content on Twitter with
 **#TidyTuesday**.

---
## Plumber

> Turn your R code into a web API

<br/>

```{r plumber-add, eval = FALSE}
#* Return the sum of two numbers
#* @param a The first number to add
#* @param b The second number to add
#* @post /sum
function(a, b) {
  as.numeric(a) + as.numeric(b)
}
```

---
class: inverse, center, middle

## Tidy Tuesday + Plumber = Tidy Tuesday API

---
class: center, middle

## What is an API?


???

machine to machine communication over the internet

show how to access TT API in browser


---
## Why?

contrib to community
data sets for non-R users
resume bullet point
fun!

---
## How

- Data sets stored in [mongo DB](https://www.mongodb.com/) database
- API fetches/sends results
- API hosted on [Digital Ocean](https://www.digitalocean.com/)

???

shout out to plumberDeploy!

---
## Mongo DB

- Document database (NoSQL)
- More flexible structure vs. tables
- Connect from R via [`{mongolite}`](https://jeroen.github.io/mongolite/)

--

.pull-left[
**SQL**

- database
- table
- record/row
]

.pull-right[
**mongo DB**

- database
- collection
- document
]


---
## API
<!-- show code -->
<!-- how does it work -->
<!-- what are the endpoints -->
<!-- etc. -->
.pull-left[
```{r avail, eval = FALSE}
#* @get /available
function(req, res, limit = 10L) {
  
  collection <- mongo(
    collection = "available",
    db = "metadata",
    url = Sys.getenv("DB_CONNECTION_URI")
  )

  available <- collection$find(
    query = '{}',
    fields = '{"_id": false}',
    sort = '{"date": -1}',
    limit = limit
  )
  collection$disconnect()

  available
}
```
]

.pull-right[
```{r data, eval = FALSE}
#* @get /data
function(req, res, date){

  fs <- gridfs(
    db = "data",
    url = Sys.getenv("DB_CONNECTION_URI")
  )
  
  
  raw_doc <- fs$read(date, progress = FALSE)
  unserialize(raw_doc$data)
  
  fs$disconnect()

  doc
}
```

]

---
## Digital Ocean + plumberDeploy


[Digital Ocean](https://www.digitalocean.com/) is a cloud provider

Databases, app platform, servers (droplets)

[plumberDeploy](https://github.com/meztez/plumberDeploy) is an R package to
 deploy your plumber API to cloud providers, from R.

<!-- what is digital ocean -->
<!-- what is a droplet -->
<!-- what is plumberDeploy -->
<!-- how does it help -->

---
class: inverse, center, middle

## Problems

problems i ran into
<!-- maybe think of a better word (issues?) -->

---
## Mongo DB

Oh the troubles i've seen...

Data too big!

spent **a lot** of time searching for a solution (Azure Cosmos DB, Amazon
 DynamoDB, ...)

Eventually landed on mongodb **gridFS**


???

mongodb document size limit: 16MB

---
## API

cors (is it ok to set it to *?)

how to handle db connections? (one connection? overhead of connect/disconnect in request)

mongolite (user permissions in db b/c C library was trying to create index, but need
 write access to make index)

---
## Digital Ocean/Deployment

linux system errors when provisioning server (tried again next day and it worked!)

.Renviron didn't deploy with app (had to ssh in and create)

deployed api but test api was already on that port so problems

how to get a domain name? what is DNS? CNAME? certificate?

<!-- plumberDeploy -->
<!-- what are some of the issues I came across? -->
<!-- how did i overcome them? -->

---
class: inverse, center, middle
## title to break things up

---
## What I learned

stress: i didn't know much going into this. only that plumber looked simple
 enough and that it was probably possible

mongodb, document databases
plumber
deploying
digital ocean
domain name

---
## Why I'm sharing

maybe you want to make a plumber api
show that it's not all rainbows and unicorns
hopefully help motivate to do side projects, especially for newer folks
yes, it may be hard, but that doesn't mean you shouldn't try

---
## Resources

links to tidy tuesday
links to tt api
links to some docs (i.e. plumber, plumberDeploy)
link to this talk (have at beginning too. title slide probably)

---
class: center, middle

# Thanks!

Slides created via the R packages:

[**xaringan**](https://github.com/yihui/xaringan)<br>
[gadenbuie/xaringanthemer](https://github.com/gadenbuie/xaringanthemer)

The chakra comes from [remark.js](https://remarkjs.com), [**knitr**](http://yihui.name/knitr), and [R Markdown](https://rmarkdown.rstudio.com).
