---
title: "Tidy Tuesday API"
subtitle: "Build and Deploy a Plumber API"  
author: "Andrew Bates"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    seal: false
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

class: center, middle

<!-- putting title slide below code chunks was causing the 'class: center' to show -->
<!--  up on the slide so put it here -->

## Tidy Tuesday API

### Andrew Bates

Slides on GitHub: `asbates/talks/tidytuesdayapi`

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo(primary_color = "#1F4257", secondary_color = "#F97B64")
```

```{css, echo = FALSE}
/* don't show slide numbers */
.remark-slide-number {
  opacity: 0;
}
```


---

## Tidy Tuesday Project

> A weekly data project aimed at the R ecosystem. 

<br/>
<br/>

Run by **Tom Mock** and the **R4DS Online Learning Community**

<br/>

A new data set is published every week. Post content on Twitter with
 **#TidyTuesday**.

---
## Plumber

> Turn your R code into a web API

<br/>

```{r plumber-add, eval = FALSE}
#* Return the sum of two numbers
#* @param a The first number to add
#* @param b The second number to add
#* @post /sum
function(a, b) {
  as.numeric(a) + as.numeric(b)
}
```

---
class: center, middle

## Tidy Tuesday + Plumber = Tidy Tuesday API

---
class: center, middle

## What is an API?


???

machine to machine communication over the internet

show how to access TT API in browser


---
## Why?

Contribute to community (R, Python, JavaScript, ...)

Resume bullet point

Fun!

---
## How

Data sets stored in [mongo DB](https://www.mongodb.com/) database

API fetches/sends results

API hosted on [Digital Ocean](https://www.digitalocean.com/)

???

shout out to plumberDeploy!

---
## Mongo DB

Document database (NoSQL)

More flexible structure vs. tables

Connect from R via [`{mongolite}`](https://jeroen.github.io/mongolite/)

--

.pull-left[
**SQL**

- database
- table
- record
]

.pull-right[
**mongo DB**

- database
- collection
- document
]


---
## API

.pull-left[
```{r avail, eval = FALSE}
#* @get /available
function(req, res, limit = 10L) {
  
  collection <- mongo(
    collection = "available",
    db = "metadata",
    url = Sys.getenv("DB_CONNECTION_URI")
  )

  available <- collection$find(
    query = '{}',
    fields = '{"_id": false}',
    sort = '{"date": -1}',
    limit = limit
  )
  collection$disconnect()

  available
}
```
]

.pull-right[
```{r data, eval = FALSE}
#* @get /data
function(req, res, date){

  fs <- gridfs(
    db = "data",
    url = Sys.getenv("DB_CONNECTION_URI")
  )
  
  
  raw_doc <- fs$read(date, progress = FALSE)
  unserialize(raw_doc$data)
  
  fs$disconnect()

  doc
}
```

]

---
## Digital Ocean + plumberDeploy


[Digital Ocean](https://www.digitalocean.com/) is a cloud provider

[plumberDeploy](https://github.com/meztez/plumberDeploy) is an R package to
 deploy your plumber API to cloud providers, from R.

???

Digital Ocean: database, app platform, servers

---
class: inverse, center, middle

## Issues


---
## Mongo DB

Oh the troubles i've seen...

Data too big!

Spent **a lot** of time searching for a solution (Azure Cosmos DB, Amazon
 DynamoDB, ...)

Eventually landed on mongodb **gridFS**


???

mongodb document size limit: 16MB

---
## API

CORS

How to handle database connections?

Permissions issues with `mongolite`

 
???

CORS: can I set it to '*'? what kind of risk?

DB connections: one connection? one connection for each request? will that be
  too much overhead?
  
mongolite: C library trying to create index, but need write access to add 
 index => had to adjust permissions for server NOT EASY TO DEBUG


---
## Digital Ocean/Deployment

Linux system errors

`.Renviron` didn't deploy with app


Domain name: `r emo::ji("shrug")`


???

Linux system: when first provisioned server, got errors I didn't understand.
 tried again next day and it worked ?

 
domain name: How to get one? What is DNS? CNAME? A certificate?

---
class: inverse, center, middle
## YOU CAN DO IT!

???

stress: I struggled *a lot*. But I learned *a lot*. You're going to run into
 issues, but that's all part of it.
 
I knew next to nothing going in. I learned about

- mongodb
- plumber
- digital ocean
- domain names
- ...


---
class: center, middle

# Thanks!

Slides created via the R packages:

[**xaringan**](https://github.com/yihui/xaringan)<br>
[gadenbuie/xaringanthemer](https://github.com/gadenbuie/xaringanthemer)

The chakra comes from [remark.js](https://remarkjs.com), [**knitr**](http://yihui.name/knitr), and [R Markdown](https://rmarkdown.rstudio.com).
